---
title: "data_preparation_script"
author: "Darren Colby"
date: "2/17/2021"
purpose: "To clean and recode the Northern Ireland Loyalty Study data so that
   it can be merged with xSub data and be analyzed."
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Standard imports and set working directory

```{r Import and set working directory}
library(asciiSetupReader)
library(dplyr)
library(data.table)
library(tidyverse)

setwd("Z:/Social Networks Northern Ireland/Data to Merge")
```

## This function reads in the original ASCII coded data, converts it to a table, 
## removes unnecessary columns, converts the table to a data frame, and selects
## only Catholic respondents, which form the pool of recruits for the IRA.
## @Parameters: The original ASCII data and a SAS script defining the columns
## @Returns: A dataframe with only the necessary columns

```{r Read survey sata from the ASCII file, echo=FALSE}
read_survey <- function(ascii_data, sas_script){
  
  loyalty <- sas_ascii_reader(ascii_data, sas_script)
  
  loyalty <- as.data.frame(loyalty[ ,c(10, 13, 78:81, 202, 376, 386)])

  return(loyalty)
}

survey_data <- read_survey("survey_data.txt", "sas_setup_script.sas")

head(survey_data)
```

## This function updates the column names to be shorter and more logical.
## @Parameters: A survey dataframe.
## @Returns: A copy of the survey dataframe with updated column names.

```{r Rename the columns}
rename_survey_columns <- function(clean_survey){
  
  names_vector <- c("hometownRel", "ukConstituency", "orgType", "orgType2", 
                    "orgType3", "orgType4", "governanceApproval",
                    "difficultFindJob", "stdLiving")
  
  updated_survey <- clean_survey
  
  colnames(updated_survey) <- names_vector
  
  return(updated_survey)
}

survey_with_new_cols <- rename_survey_columns(survey_data)

head(survey_with_new_cols)
```

## This (extremely long) function recodes the character columns in the NI 
## Loyalty Study into factor, binary, and numeric variables in accordance with 
## the codebook.
## @Parameters: survey, the survey used in this study.
## @Returns: A recoded survey.

```{r Recode the variable levels}
recode_survey <- function(survey){
  
  agreement <- c("DISAGREE" = "1", "NO OPINION"= "2", "AGREE" = "3", 
                 .default = NA_character_)
  
  yes_no <- c("NO" = "1", "YES" = "3", "DEPENDS" = "2", 
              .default = NA_character_)
  
  survey_new_code <- survey_with_new_cols
  
  survey_new_code$hometownRel <- recode_factor(
    survey_new_code$hometownRel,
    "LIVES IN ROMAN CATHOLIC AREA PSU" = "1",
    "LIVES IN MIXED AREA PSU" = "2",
    "LIVES IN PROTESTANT AREA PSU" = "3",
    .default = NA_character_
  )
  
  survey_new_code$governanceApproval <- recode_factor(
    survey_new_code$governanceApproval,
    "ASHAMED" = "1",
    "NOT VERY PROUD" = "2",
    "FAIRLY PROUD" = "3",
    "VERY PROUD" = "4",
    .default = NA_character_
  )
  
  survey_new_code$difficultFindJob <- recode_factor(
    survey_new_code$difficultFindJob,
    !!!yes_no
  )
  
  survey_new_code$stdLiving <- recode_factor(
    survey_new_code$stdLiving,
    "POOR" = "1",
    "NOT VERY SATISFACTORY" = "2",
    "FAIRLY SATISFACTORY" = "3",
    "VERY SATISFACTORY" = "4",
    .default = NA_character_
  )
  
  survey_new_code$ukConstituency <- recode_factor(
    survey_with_new_cols$ukConstituency,
    "ARMAGH" = "Armagh",
    "BELFAST EAST" = "Belfast East",
    "BELFAST NORTH" = "Belfast North",             
    "NA" = "Belfast South",
    "BELFAST WEST" = "Belfast West",
    "FERMANAGH AND SOUTH TYRONE" = "Fermanagh and South Tyrone",
    "LONDONDERRY" = "Londonderry",                
    "MID-ULSTER" = "Mid Ulster",                 
    "NORTH ANTRIM" = "North Antrim",              
    "NORTH DOWN" = "North Down",                 
    "SOUTH ANTRIM" = "South Antrim",               
    "SOUTH DOWN" = "South Down"
  )
  
  return(survey_new_code)
}

recoded_survey <- recode_survey(survey_with_new_cols)

head(recoded_survey)
```

## This function adds up all of the unique organizations for each constituency 
## across the orgType, orgTYpe2, orgType3, and orgType 4 columns.
## @Parameters: A survey dataframe.
## @Returns: A vector of the number of unique organizations in each constituency

```{r Add up the organizations in each constituency}
aggregate_organizations <- function(data){
  
  org_count <- tapply(data$orgType, data$ukConstituency, unique)
  
  org_count2 <- tapply(data$orgType2, data$ukConstituency, unique)
  
  org_count3 <- tapply(data$orgType3, data$ukConstituency, unique)
  
  org_count4 <- tapply(data$orgType4, data$ukConstituency, unique)
  
  flat_list <- mapply(c, org_count, org_count2, org_count3, org_count4, 
                      SIMPLIFY = FALSE)
  
  flat_list <- lapply(flat_list, unique)
  
  flat_list <- lapply(flat_list, function(glist) 
    glist[!grepl("NA <OR INAP> NO FURTHER MENTIONS", glist)])
  
  flat_list <- lapply(flat_list, length)
  
  return(as.vector(unlist(flat_list)))
}
```

## Make columns numeric so they can be passed to the group_by function

```{r Convert survey columns to numeric}
recoded_survey$hometownRel <- as.numeric(recoded_survey$hometownRel)
recoded_survey$governanceApproval <- as.numeric(
   recoded_survey$governanceApproval)
recoded_survey$stdLiving <- as.numeric(recoded_survey$stdLiving)
recoded_survey$difficultFindJob <- as.numeric(recoded_survey$difficultFindJob)
```


## This function aggregates survey data by UK consituency.
## @Parameters: survey_data, a dataframe.
## @Returns: A dataframe of aggregated survey data.

```{r Aggregate the survey data for each constituency}
aggregate_survey_data <- function(survey_data){
  
  orgs <- aggregate_organizations(survey_data)
  
  survey <- survey_data %>%
     group_by(ukConstituency) %>%
     summarise_if(is.numeric, mean, na.rm = TRUE) %>%
     cbind(orgs)
  
  return(survey)
}

aggregated_survey <- aggregate_survey_data(recoded_survey)
     
head(aggregated_survey)
```

## This function reads in the event data, and filters it for the relevant columns.
## @Parameters: The name of a csv file of event data.
## @Returns: Filtered event data

```{r Get clean event data}
get_clean_event_data <- function(filename){
  
  event_data <- read.csv(filename)
  
  event_data <- event_data %>%
    relocate(id, lat, long, date, name, areaSqKm, boundary)
  
  return(event_data)
}

event_data <- get_clean_event_data("xSub.csv")

head(event_data)
```


## Merge the datasets by constituency


```{r Merge the datasets}
final_data <- merge(event_data, aggregated_survey, by.x = "name", 
                      by.y = "ukConstituency") %>%
                      relocate(id)

summary(final_data)
```

## Add a column for the year and the number of years since 1969

```{r}
final_data$year <- as.numeric(substr(final_data$date, 1, 
                                     nchar(final_data$date)-4))

final_data$yearsSince1969 <- final_data$year - 1969
```

## Add a column for government attacks per year in each constituency

```{r}
  
  attacks <- final_data %>%
    group_by(name, year) %>%
    summarise(events = sum(initGovt))

  attacks <- attacks %>%
    arrange(year)

  Armagh <- attacks %>%
    filter(name == "Armagh") %>%
    mutate(lag = lag(events))

  belfastEast <- attacks %>%
    filter(name == "Belfast East") %>%
    mutate(lag = lag(events))

  belfastNorth <- attacks %>%
    filter(name == "Belfast North") %>%
    mutate(lag = lag(events))

  belfastWest <- attacks %>%
    filter(name == "Belfast West") %>%
    mutate(lag = lag(events))

  Fermanagh <- attacks %>%
    filter(name == "Fermanagh and South Tyrone") %>%
    mutate(lag = lag(events))

  Londonderry <- attacks %>%
    filter(name == "Londonderry") %>%
    mutate(lag = lag(events))

  midUlster <- attacks %>%
    filter(name == "Mid Ulster") %>%
    mutate(lag = lag(events))

  northAntrim <- attacks %>%
    filter(name == "North Antrim") %>%
    mutate(lag = lag(events))

  southAntrim <- attacks %>%
    filter(name == "South Antrim") %>%
    mutate(lag = lag(events))

  northDown <- attacks %>%
    filter(name == "North Down") %>%
    mutate(lag = lag(events))

  belfastSouth <- attacks %>%
    filter(name == "Belfast South") %>%
    mutate(lag = lag(events))

  southDown <- attacks %>%
    filter(name == "South Down") %>%
    mutate(lag = lag(events))

  attacks <- bind_rows(Armagh, belfastEast, belfastNorth, belfastWest, Fermanagh,
                      Londonderry, midUlster, northAntrim, southAntrim, 
                      northDown, belfastSouth, southDown) %>%
                      select(-events)

  final_data <- merge(final_data, attacks, by = c("name", "year")) 
  
  names(final_data)[names(final_data) == "lag"] <- "govtAttacks"
  

```

## Add a column for rebel attacks per year in each constituency
```{r}
  attacks <- final_data %>%
    group_by(name, year) %>%
    summarise(events = sum(initRebel))

  attacks <- attacks %>%
    arrange(year)

  Armagh <- attacks %>%
    filter(name == "Armagh") %>%
    mutate(lag = lag(events))

  belfastEast <- attacks %>%
    filter(name == "Belfast East") %>%
    mutate(lag = lag(events))

  belfastNorth <- attacks %>%
    filter(name == "Belfast North") %>%
    mutate(lag = lag(events))

  belfastWest <- attacks %>%
    filter(name == "Belfast West") %>%
    mutate(lag = lag(events))

  Fermanagh <- attacks %>%
    filter(name == "Fermanagh and South Tyrone") %>%
    mutate(lag = lag(events))

  Londonderry <- attacks %>%
    filter(name == "Londonderry") %>%
    mutate(lag = lag(events))

  midUlster <- attacks %>%
    filter(name == "Mid Ulster") %>%
    mutate(lag = lag(events))

  northAntrim <- attacks %>%
    filter(name == "North Antrim") %>%
    mutate(lag = lag(events))

  southAntrim <- attacks %>%
    filter(name == "South Antrim") %>%
    mutate(lag = lag(events))

  northDown <- attacks %>%
    filter(name == "North Down") %>%
    mutate(lag = lag(events))

  belfastSouth <- attacks %>%
    filter(name == "Belfast South") %>%
    mutate(lag = lag(events))

  southDown <- attacks %>%
    filter(name == "South Down") %>%
    mutate(lag = lag(events))

  attacks <- bind_rows(Armagh, belfastEast, belfastNorth, belfastWest, Fermanagh,
                      Londonderry, midUlster, northAntrim, southAntrim, 
                      northDown, belfastSouth, southDown) %>%
                      select(-events)

  final_data <- merge(final_data, attacks, by = c("name", "year"))
```

## Change the column names

```{r}
new_names <- c("name", "year", "id", "lat", "long", "date", "areaSqKm", 
               "boundary", "initGovt", "initRebel", "initCiv", "initOther",
               "targetGovt", "targetRebel", "targetCiv", "targetOther", 
               "direct", "religion", "governance", "employment", "stdLiving",
               "clubs", "yearsSince1969", "govtAttacks", "rebelAttacks")

colnames(final_data) <- new_names
```



## Finally, write the data to a csv file

```{r Write the data to a csv file}
write.csv(final_data, file = 'finalData.csv')
```

